{Gravitational wave sources in our Galactic backyard:}\\{Predictions for BHBH, BHNS and NSNS binaries detectable with LISA

### 2022-01-22 paper/tex/sections/introduction.tex
clarify where stellar-mass sources come from
remove sentence about BHNS counterparts

The Laser Interferometer Space Antenna (LISA, \citealp{Amaro-Seoane+2017, Colpi+2019}) will provide observations in an entirely new regime of gravitational waves. LISA will observe at lower frequencies ($10^{-5} \lesssim f / \unit{Hz} \lesssim 10^{-1}$) than ground-based detectors and so will enable the study of sources that are imperceptible by ground-based detectors, such as the mergers of supermassive black holes and extreme mass-ratio inspirals \citep[e.g.][]{Begelman+1980, Klein+2016}. Moreover, this frequency regime is also of interest for the detection of \textit{local} stellar-mass double compact objects (DCOs) millions of years before their merger. This presents an opportunity for both multi-messenger detections to search for electromagnetic counterparts, as well as multiband gravitational-wave detections that can help to constrain binary characteristics \citep[e.g.][]{Sesana+2016, Gerosa+2019}. In addition, LISA will be able to measure the eccentricities of DCOs, which may yield further constraints on binary evolution, differentiate between formation channels and distinguish between DCO types \citep[e.g.][]{Nelemans+2001, Breivik+2016, Antonini+2017, Rodriguez+2018}. Unlike ground-based detectors, LISA only detects stellar-mass sources in local galaxies, with the majority residing in the Milky Way. These sources could be used as a probe for the structure of our galaxy \citep[e.g.][]{Korol+2019}.

@@

One particularly interesting class of potential LISA sources is BHNSs. With the recent detection of two BHNSs by the LIGO scientific collaboration, the existence of these DCOs has been confirmed \citep{TheLIGOScientificCollaboration+2021}. However, with only two detections (not including the low-confidence candidate GW190426, \citealt{Abbott+2020_GWTC2}, or GW190425, GW190814 and GW190917 which have not been ruled out as BHNSs, \citealt{Abbott+2020_GW190425,Abbott+2020_GW190814, GWTC_2_1}) and no electromagnetic counterparts, the formation rate and properties of BHNSs are still uncertain. Current predictions for the merger rate of BHNSs range across three orders of magnitude \citep[e.g.][]{Abadie+2010, Mandel+2021} so the number of detections in LISA will be important in reducing this uncertainty, thereby refining our understanding of the remnants and evolution of massive stars. Similar to NSNSs, these binaries are also expected to have electromagnetic counterparts. A distinctly exciting possibility is the detection of a pulsar--BH system or millisecond pulsar--BH system \citep{Narayan+1991, Pol+2021}. These systems could be observed not only by LISA, but also radio telescopes such as MeerKAT and the Square Kilometre Array (SKA, \citealt{Dewdney+2009}), which would help to improve the measurement of individual system parameters and to constrain uncertain binary evolution processes \citep[e.g.][]{Pfahl+2005,Chattopadhyay+2020}.

>>>>

The Laser Interferometer Space Antenna (LISA, \citealp{Amaro-Seoane+2017, Colpi+2019}) will provide observations in an entirely new regime of gravitational waves. LISA will observe at lower frequencies ($10^{-5} \lesssim f / \unit{Hz} \lesssim 10^{-1}$) than ground-based detectors and so will enable the study of sources that are imperceptible by ground-based detectors, such as the mergers of supermassive black holes and extreme mass-ratio inspirals \citep[e.g.][]{Begelman+1980, Klein+2016}. Moreover, this frequency regime is also of interest for the detection of \textit{local} stellar-mass double compact objects (DCOs) millions of years before their merger. This presents an opportunity for both multi-messenger detections to search for electromagnetic counterparts, as well as multiband gravitational-wave detections that can help to constrain binary characteristics \citep[e.g.][]{Sesana+2016, Gerosa+2019}. In addition, LISA will be able to measure the eccentricities of DCOs, which may yield further constraints on binary evolution, differentiate between formation channels and distinguish between DCO types \citep[e.g.][]{Nelemans+2001, Breivik+2016, Antonini+2017, Rodriguez+2018}. Unlike ground-based detectors, \referee{the majority of stellar-mass sources detected in LISA will reside in local galaxies, with most situated in the Milky Way}. These sources could be used as a probe for the structure of our galaxy \citep[e.g.][]{Korol+2019}.

@@

One particularly interesting class of potential LISA sources is BHNSs. With the recent detection of two BHNSs by the LIGO scientific collaboration, the existence of these DCOs has been confirmed \citep{TheLIGOScientificCollaboration+2021}. However, with only two detections (not including the low-confidence candidate GW190426, \citealt{Abbott+2020_GWTC2}, or GW190425, GW190814 and GW190917 which have not been ruled out as BHNSs, \citealt{Abbott+2020_GW190425,Abbott+2020_GW190814, GWTC_2_1}) and no electromagnetic counterparts, the formation rate and properties of BHNSs are still uncertain. Current predictions for the merger rate of BHNSs range across three orders of magnitude \citep[e.g.][]{Abadie+2010, Mandel+2021} so the number of detections in LISA will be important in reducing this uncertainty, thereby refining our understanding of the remnants and evolution of massive stars. A distinctly exciting possibility for BHNSs is the detection of a pulsar--BH system or millisecond pulsar--BH system \citep{Narayan+1991, Pol+2021}. These systems could be observed not only by LISA, but also radio telescopes such as MeerKAT and the Square Kilometre Array (SKA, \citealt{Dewdney+2009}), which would help to improve the measurement of individual system parameters and to constrain uncertain binary evolution processes \citep[e.g.][]{Pfahl+2005,Chattopadhyay+2020}.
